The markings of the trade: exploring the patterns of olive oil production in Roman Baetica

### 2020-05-06 Paper/jaccardbaetica.tex
más comentarios

%\xavi{este aim es demasiado general}
%This paper aims to understand the production dynamics in connection with different areas within the Roman Empire. 
This paper aims to study the olive oil market connection between province by calculating the similarity index. Specifically, our work pretends to detect microeconomic processes focused on a commercial product from a specific province \citep{isaksen_network_2006}. We want to understand the pattern of olive oil production linked to amphora workshops and amphoric stamps used to mark them. We focus here on exploring the economic relation between stamps and amphora production and distribution centres.

@@

Some stamps show a more specific chronology while the majority of them display a large activity of production that it can be difficult for specifying an accurate chronology. This could be due to two reasons; firstly, most of the workshops were partially excavated and focused on archaeological surveys in order to collect the maximum stamps as possible; secondly, Dressel 20 was produced during almost three centuries with apparently few changes \citep{berni_piero_chapter_2017}.

\section{Britannia and Germania}

@@

It is well known that this product was not frequently consumed by indigenous population. The absence of olive oil importation before the conquest can be reflected by the shortage of this product until the arrival of a high percentage of Dressel 20 amphorae to \textit{Britannia} \citep[ 1]{carreras_britannia_1998}. Additionally, the land of \textit{Britannia} showed inadequate for the olive oil production due to low environmental conditions. This issue was necessary for creating an important export apparatus from \textit{Baetica} to supply Roman army.

@@

Finally, The increase in olive oil exports would experience a progressive slowdown
from the third century A.D., coinciding with the change in market strategy in the
Empire. As a result, a progressive decrease of Dressel 20 is documented being gradually replaced for Dressel 23 amphorae \citep{rodriguez1991aceite,
millet_anforas_1998}.

@@

\section{A potential indicator of Roman economy: Dressel 20 stamps}

@@

\section{Method and Analysis}

The assumption of this study is analysing the effect of the production patterns between different centres. We are specially interested in identifying links between producers and consumption centres using amphoric stamps. To do that, we use the CEIPAC database to collect the stamps from different places. The CEIPAC dataset contains over 50.000 of epigraphy records found in amphorae, mostly from \textit{Monte Testaccio}. Therefore, this study proposes a robust baseline to explore the distribution of Baetican olive oil production by computing the spatial correlation between stamps. A way to analyse is to use a quantitative framework to measure the similarity between amphoric stamps. Here we use an ecological approach based on three steps: a) to detect similarities between stamp codes, b) to explore a potentially spatial correlation and c) to establish a correlation between similarity of stamps and spatial distance. 
%\xavi{no veo la parte evolutiva en el paper; por otra parte, si es relevante no la lanzaría aquí al vuelo; hace falta explicar por qué es necesaria, qué implica, etc.}

@@

@@

@@

@@

@@

@@

\subsection{Reception centres: Britannia and Germania}

>>>>

%\xavi{este aim es demasiado general}
%This paper aims to understand the production dynamics in connection with different areas within the Roman Empire.

\xavi{le das demasiada importancia al similarity index. Es una simple herramienta de tu trabajo, y no un objetivo}.
This paper aims to study the olive oil market connection between province by calculating the similarity index. Specifically, our work pretends to detect microeconomic processes focused on a commercial product from a specific province \citep{isaksen_network_2006}. We want to understand the pattern of olive oil production linked to amphora workshops and amphoric stamps used to mark them. We focus here on exploring the economic relation between stamps and amphora production and distribution centres.

@@

\xavi{Estás segura de esto? Porque aquí hay 2 cosas: 1. datar bien y 2 . período de uso. Yo no tengo claro que la mayoría de sellos se usen durante mucho tiempo. En qué está basada la idea? Deberías detallar un poco más este temita cronológico y explicar como sabemos el período de uso.}
Some stamps show a more specific chronology while the majority of them display a large activity of production that it can be difficult for specifying an accurate chronology. This could be due to two reasons; firstly, most of the workshops were partially excavated and focused on archaeological surveys in order to collect the maximum stamps as possible; secondly, Dressel 20 was produced during almost three centuries with apparently few changes \citep{berni_piero_chapter_2017}.

\section{Britannia and Germania}
\xavi{este título de sección debería cambiarse por algo más descriptivo. Case studies o algo asínnn? Por otra parte, igual podrías dividirlo en 2 subsections}

\xavi{en general, echo a faltar un poco más de detalle en esta discusión y en los datos que tienes; en qué se parecen y diferencian las 2 provincias? Si hablas del uso por las campañas o las guarniciones, donde están? Qué strategic points de redistribución hay? Tampoco a saco, pero entender mejor cómo llega el aceite, cómo se redistribuye y quién y dónde se consume}

@@

It is well known that this product was not frequently consumed by indigenous population \xavi{CITATION NEEDED}. The absence of olive oil importation before the conquest can be reflected by the shortage of this product until the arrival of a high percentage of Dressel 20 amphorae to \textit{Britannia} \citep[ 1]{carreras_britannia_1998}. Additionally, the land of \textit{Britannia} showed inadequate for the olive oil production due to low environmental conditions. This issue was necessary for creating an important export apparatus from \textit{Baetica} to supply Roman army.

@@

Finally, The increase in olive oil exports would experience a progressive slowdown from the third century A.D., coinciding with the change in market strategy in the Empire. As a result, a progressive decrease of Dressel 20 is documented being gradually replaced for Dressel 23 amphorae \citep{rodriguez1991aceite,millet_anforas_1998}.

@@

\section{A potential indicator of Roman economy: Dressel 20 stamps}
\xavi{no sería mejor que esta sección esté antes de hablar de los casos de estudios? Así, podrías ya hablar de sellos encontrados en germania y britannia}

@@

\section{Method and Analysis}
\xavi{estas 2 cosas nunca van juntas. En Method explicas las técnicas, y Analysis en general no se usa porque es el todo del trabajo}

The assumption of this study is analysing the effect of the production patterns between different centres \xavi{esto no es ninguna asunción no??}. We are specially interested in identifying links between producers and consumption centres using amphoric stamps. To do that, we use the CEIPAC database to collect the stamps from different places. The CEIPAC dataset contains over 50.000 of epigraphy records found in amphorae, mostly from \textit{Monte Testaccio}. Therefore, this study proposes a robust baseline to explore the distribution of Baetican olive oil production by computing the spatial correlation between stamps. A way to analyse is to use a quantitative framework to measure the similarity between amphoric stamps. Here we use an ecological approach based on three steps: a) to detect similarities between stamp codes, b) to explore a potentially spatial correlation and c) to establish a correlation between similarity of stamps and spatial distance. 
%\xavi{no veo la parte evolutiva en el paper; por otra parte, si es relevante no la lanzaría aquí al vuelo; hace falta explicar por qué es necesaria, qué implica, etc.}

@@

\xavi{que me quede claro...la mayoría de workshops solo tienen 1 sello??? Y hay un workshop con como 700 sellos distintos??? }

@@

\xavi{si el eje es logarítmico deberías mencionarlo}

@@

\xavi{vale pero qué es cada punto? un workshop?}

@@

\xavi{Esta estructura no me convence...tenemos que pensar algo mejor, porque ahora es: 1. hablo de britania/germania, 2. hablo de sellos en general, 3. hablo de sellos en britania/germania. Quizás juntar esto con lo de los casos de estudio, o quizás hablar de los 2 casos de estudio en la intro, y dividir la sección anterior en 2 (lo más general a la intro, lo más concreto aquí)}

@@

\xavi{en stamps o en yacimientos? o en provincias?}

@@

\subsection{Reception centres: Britannia and Germania}
\xavi{mejor consumption?}
